---
title: "p8105_hw1_bw2595"
author: "Blair Wigsten"
date: "9/17/2019"
output: github_document
---

Problem 1
```{r hw1_q1}

library(tidyverse)

hw1q1_df = tibble(
  norm_samp = rnorm(8),
  vec_logical = c(norm_samp > 0),
  vec_char = c("A", "B", "C", "D", "E", "F", "G", "H"),
  vec_fac = factor(c("low", "med", "high", "low", "med", "high", "low", "med"))
)

mean(pull(hw1q1_df, norm_samp))
mean(pull(hw1q1_df, vec_logical))
mean(pull(hw1q1_df, vec_char))
mean(pull(hw1q1_df, vec_fac))
```

Only the first 2 variables have a mean value outputted. We cannot take the mean of a character or factor variable, so we get an error that the argument is not numeric or logical.  

```{r hw1q1 include = FALSE}

as.numeric(pull(hw1q1_df, vec_logical))
as.numeric(pull(hw1q1_df, vec_char))
as.numeric(pull(hw1q1_df, vec_fac))

```
When we do this, we get number values for our logical and factor variables, so we could take the mean here. But since the character vector does not connect to numbers, we cannot convert to numeric and are given the "NAs introduced by coercion" error. 
















Problem 2
```{r Problem_2a}

hw1q2_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  vec_log = c(x + y > 1), 
  vec_num = as.numeric(vec_log),
  vec_fac = as.factor(vec_log)
)
```

The size of the dataset and number of columns is
`r nrow(hw1q2_df) ncol(hw1q2_df)`.

```{r}
ggplot(plot_df, aes(x = x, y = y), color = ...) + geom_point()
```

