---
title: "p8105_hw1_bw2595"
author: "Blair Wigsten"
date: "9/17/2019"
output: github_document
---

# Problem 1

## Load tidyverse and create the logical, character, and factor vectors within the dataframe
* a random sample of size 8 from a standard Normal distribution
* a logical vector indicating whether elements of the sample are greater than 0
* a character vector of length 8
* a factor vector of length 8, with 3 different factor “levels”
```{r hw1q1a}
library(tidyverse)

set.seed(100)

hw1q1_df = tibble(
  norm_samp = rnorm(8),
  vec_logical = c(norm_samp > 0),
  vec_char = c("A", "B", "C", "D", "E", "F", "G", "H"),
  vec_fac = factor(c("low", "med", "high", "low", "med", "high", "low", "med"))
)
```
* This creates a dataframe called hw1q1_df comprised of a random sample of size 8 from a standard normal distribution (norm_samp), a logical vector indicating whether elements of the sample are greater than 0 (vec_logical), a character vector of length 8 (vec_char), and a factor vector of length 8, with 3 different levels (vec_fac).

## Try to take the mean of each variable in the dataframe
```{r}
mean(pull(hw1q1_df, norm_samp))
mean(pull(hw1q1_df, vec_logical))
mean(pull(hw1q1_df, vec_char))
mean(pull(hw1q1_df, vec_fac))
```
## What works and what doesn't?
* Only the first 2 variables have a mean value outputted. We cannot take the mean of a character or factor variable, so we get an error that the argument is not numeric or logical for the second two lines of code.  

## Apply as.numeric function to our logical, character, and factor vectors
```{r eval = FALSE}
as.numeric(pull(hw1q1_df, vec_logical))
as.numeric(pull(hw1q1_df, vec_char))
as.numeric(pull(hw1q1_df, vec_fac))
```
* When we apply the as.numeric function to our vectors, we get number values for our logical and factor variables, so we can take their means now. However, since the character vector does not connect to numbers, we cannot convert to numeric and are given the "NAs introduced by coercion" error. 

## Converting vectors 
* convert the logical vector to numeric, and multiply the random sample by the result
* convert the logical vector to a factor, and multiply the random sample by the result
* convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result
```{r hw1q1c}
as.numeric(pull(hw1q1_df, vec_logical)) * pull(hw1q1_df, norm_samp)

as.factor(pull(hw1q1_df, vec_logical)) * pull(hw1q1_df, norm_samp)

as.numeric(as.factor(pull(hw1q1_df, vec_logical)))*pull(hw1q1_df, norm_samp)
```
* The first line of code converts our logical vector to numeric, and multiplies the random sample by the result. However, you cannot multiply a factor vector, so we get an error in the second line of code that the symbol * is not meaningful for factor vectors. The third line of code converts the logical vector to a factor vector, then converts the result to numeric, and multiplies that result by the random sample.

# Problem 2

## Creating a dataframe that consists of 
* x: a random sample of size 500 from a standard Normal distribution
* y: a random sample of size 500 from a standard Normal distribution
* A logical vector indicating whether x + y > 1
* A numeric vector created by coercing the above logical vector
* A factor vector created by coercing the above logical vector
```{r Problem_2a}

set.seed(100)

hw1q2_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  vec_log = c(x + y > 1), 
  vec_num = as.numeric(vec_log),
  vec_fac = as.factor(vec_log)
)
```
* This creates a dataframe comprised of a random sample of size 500 from a standard normal distribution (x), a random sample of size 500 from a standard normal distribution (y), a logical vector indicating whether x + y > 1 (vec_log), a numeric vector created by coercing vec_log (vec_num), and a factor vector created by coercing vec_log (vec_fac).

## Inline code describing the data
There are `r nrow(hw1q2_df)` rows in the dataframe and `r ncol(hw1q2_df)` columns. The mean value of x is `r mean(pull(hw1q2_df, x))`, the median value of x is `r median(pull(hw1q2_df, x))`, and the standard deviation of x is `r sd(pull(hw1q2_df, x))`. The proportion of cases for which x + y > 1 is `r sum(pull(hw1q2_df, vec_log))/500`

## Create scatterplots of logical, numeric, and factor vectors
```{r Problem2_scatterplot}
ggplot(hw1q2_df, aes(x = x, y = y, color = vec_log)) + geom_point()
ggsave("scatter_plot.pdf", height = 6, width = 6)

ggplot(hw1q2_df, aes(x = x, y = y, color = vec_num)) + geom_point()
ggplot(hw1q2_df, aes(x = x, y = y, color = vec_fac)) + geom_point()
```
* The first plot of y vs. x using the logical variable has two different colors for points. The red points indicate that the observation does not satisfy the statement that the sum of x and y is greater than 1, and the blue dots satisfy that condition. 
* The second plot of y vs. x using the numeric vector has a color scale, if we were to have points that were not either 0 or 1. Since we are converting the logical vector to numeric, we will only have 0 or 1 values, so that is why we only see 2 distinct colors, rather than the gradient shown in the key. 
* The third plot of y vs. x using the factor vector is similar to the logical vector graph. The red points indicate that the observation does not satisfy the logical vector statement, while the blue ones show points that satisfy the condition that     x + y > 1. 